---
title: "loading well data"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## LAS files

The LAS format is widely used in the Oil and Gas industry and is short from Log ASCII Standard, in this notebook we show a short workflow to load and qc multiple files. I´m using some files donated by [Geolink](https://www.geolink-s2.com/) to the geoscience community.

<br>

## The libraries

I will use to libraries:

-   the [tidyverse](https://www.tidyverse.org/) to perform data wrangling and plotting,

-   [petroreadr](https://github.com/ravenroadresources/petroreadr) from **Ravenroadresources** to load the **LAS** files.

-   [skimr](https://docs.ropensci.org/skimr/), to get an exccelent summary of the data. 

<br>


```{r echo=TRUE, message=FALSE, warning=FALSE}

library(tidyverse)
library(petroreadr)
library(skimr)


```

<br>

## Loading the files

The library indicates in the documentation a few options, I normally choose:

-   load directly to a dataframe, which is my go to mode.\
-   set the verbose option to `TRUE`, this is particularly useful when working with large amount of wells/logs, to have the peace of mind that the machine is working and has not hanged.


<br>

```{r echo=TRUE, message=FALSE, warning=FALSE}

pathname <- "../../data/GEOLINK_Lithology and wells NORTH SEA/"

lasfiles <- list.files(pathname)
lasfiles <- lasfiles[grepl(".las", lasfiles)]
 
df<-read_las(file.path(pathname, lasfiles), verbose = TRUE)$data

  
  
```

<br>

## the loaded data

We have obtained a dataframe with all logs, and a column with the well name that makes our life easier for plotting, filtering, etc.


<br>

```{r}
skim(df)
```

<br>

## Quick view of a log

to check whether a well log is present in all wells, and compare some of their stats, the box plot is a powerful tool.

<br>

```{r}

box_1 <-
  df %>%
  ggplot(aes(WELL, GR)) +
  geom_boxplot() +
  theme_bw()+
  theme(axis.text.x=element_text(angle=90,hjust=1)
        )


box_1

```

<br>

A quick view shows some negative values in well *34/10-12* and some potential outliers in well *35/9-2*, we can repeat the plot cropping on a range to have a better view.

<br>

```{r echo=FALSE, message=FALSE, warning=FALSE}

box_1 + ylim(0, 200)
  
```

<br>

We can repeat the same for all logs at once to get a quick idea , however there are at least two points to consider when doing this:

-   the size of the plot will need to be increased to properly show all logs,

-   if you want to visualize a zoom of a particular log, you will need to apply a filter to the data before appliying the pivot.

-   similarly, when dealing with logs that behave logarithmicaly, I dont´t know how to set a logarithmic scale to a particular log using faceting, therefore I apply a log transform before the pivoting step.

<br>

```{r fig.height=60, fig.width=8, message=FALSE, warning=FALSE}

df %>% 
  mutate(log10_RDEP= log10(RDEP)) %>% 
  select(-RDEP) %>% 
  filter( GR >   0 ,
         GR  < 200 )  %>% 
  pivot_longer(!c("WELL"), names_to = "logs", values_to = "value") %>% 
  drop_na(value) %>%
  ggplot()+
  geom_boxplot(aes(WELL, value))+
  facet_wrap(~logs, scales = "free", ncol = 1) +
  theme_bw()+
  theme(axis.text.x=element_text(angle=90,hjust=1))


```

<br>

## Density neutron plot

Just a quick plot set creating a color scale between 0 and 100, the values below 0 are filtered from the plot, but values above 100 are colored as 100.

<br>

```{r message=FALSE, warning=FALSE}
df %>% 
  filter(GR>0) %>% 
  ggplot(aes(RHOB,NPHI, color = GR))+
  geom_point( size = 1)+
  facet_wrap(~WELL, ncol = 2)+
  xlim(1,max(df$RHOB))+
  theme_bw()+
  scale_color_gradient(low = "yellow", high = "brown",limits = c(0,100), 
                       oob = scales::squish
                       )
```
